---
  pdf_document: default
output:
  html_document: default
pdf_document: default
title: 'Survie grande dimension'
html_notebook: default
---
```{r}

library(survival)
library(rootSolve) 
library(nleqslv)

```

## data generation

 Here, we first generate a large database "surv_data" of size $m$ containing the true event times, the censoring indicators, and the  Cox model covariates.
 
```{r}

Generate_data <- function(m,n,beta){
  ##covarites data
  X1 = rnorm(m,0,1)
  X2 = rbinom(m,size = 1, prob = 0.8)
  X  = as.matrix(cbind(X1, X2))
  
  U  = runif(m,0,1) # Time to event variable
  Tt = -log(U)/exp((X%*%beta))#lambda=1
  # Constant right censored
  #c = quantile(Tt[,1], probs = 0.75)
  c  = 1.777907
  
  Time      = pmin(Tt,c) # le vrai temps
  delta     = as.numeric(Tt<=c)
  surv_data = data.frame(Time,delta,X)
  colnames(surv_data) = c('Time','delta',paste("X", 1:2, sep = ""))
 
    return(surv_data=surv_data )
}

```


 
```{r}
m    = 500
n    = round((m*80)/100)
beta = c(0.5,-0.5)
data = Generate_data(m,n,beta)
head(data)
```


Method1 presentation: naive method.
 

- We isolate a database called "data_true" of size $n$, which contains the true information on event times and censors. Then we consider a database named "XB" of size $m$ which contains all the information on the covariates.

- Finally, since we have assumed that the records linkage probabilities $Q$ are known,  we construct a database "naive_data" of size $n< m$, which contains the information on time, censures and covariates selected with a probability $Q$ according to a multinomial distribution in order to insert matching errors.

- We estimate the beta parameter by solving the estimating equation "equa_naive" using Newton's numerical method in "coxph_equa_naive ".

```{r}

#########  multinomiale Q

Matrix_function<-function(n,m,C,beta, surv_data){

  #surv_data <- Generate_data(m,n,beta)
  X  =  surv_data[,3:(p+2)]
  XB = as.matrix(X, ncol = p) #Only covariates
  
  ###### True-link data: The first n individual from B
  data_true = surv_data[1:n,]
  data_A = surv_data[1:n,1:2] #Survival time and censored indicator
  
  
  Q = matrix(0,n,m)
  Lvec = matrix(0,n,m)
  for (i in 1:n) {
    if(i!=1 & i!=n){ 
      pro = rep(0,m)
      pro[i] = C[1] #Probability of True links
      pro[i+1]= C[3]
      pro[i-1]=C[3]
      Q[i,]<-pro
      Lvec[i,] <- rmultinom(1, size = 1, pro = pro)} 
    if(i==1)  {
      pro = rep(0,m)
      pro[i] =  C[1] #Probability of True links
      pro[i+1]= C[2]
      pro[i+2]= C[3]
      Q[i,]<-pro
      Lvec[i,] <- rmultinom(1, size = 1, pro = pro)} 
    if(i==n){ 
      pro = rep(0,m)
      pro[i] = C[1] #Probability of True links
      pro[i-1]=C[2]
      pro[i-2]=C[3]
      Q[i,]<-pro
      Lvec[i,] <- rmultinom(1, size = 1, pro= pro) }  
  }
  
  Lvec
  X_naive <-Lvec %*% XB
  colnames(X_naive)<- c("Z1","Z2")
  
  # Naive data
  data_naive <- cbind( data_A, X_naive)

  return(list( data_true=data_true, data_A= data_A,XB=XB, data_naive=data_naive,Q=Q, Lvec= Lvec))
}

```


```{r}
p = 2
C = c(0.8,0.1,0.1)
surv_data = data

mf = Matrix_function(n,m,C,beta,surv_data)
data_true = mf$data_true
data_naive = mf$data_naive

Ts = mf$data_true$Time
event = mf$data_true$delta
Q = mf$Q
XB = mf$XB
Z = as.matrix(data_naive[,3:(p+2)])

head(data_true)
head(data_naive)

```
################## ##################################################




Methode 4 presentation: Estimate method

 $\textbf{ simulation steps for parameter estimation} $

```{r}

#1. Finding the risk set R(t) given some time t

GetRiskSet <- function(time_of_interest, time_vector, event_vector) {
  
  return(which(((time_vector == time_of_interest & event_vector == 1) | (time_vector > time_of_interest))))
  
}
```


```{r}

#####
#number of observations before time t

  observe<- function(time_of_interest, time_vector) {
    return(which( (time_vector <= time_of_interest) ))
  }

```


1) Parameters initialization.

$\lambda_0^{0}=0.5 \quad \text{if} \quad \delta_i=1, \quad \text{and} \quad 0 \quad \text{otherwise}; \quad \beta^{0}=(0.1,0.1)$



2) Basic cumulative function is represented by the function "Funct_lambda2"


   $ \Lambda_0^{0}(t_i) =  \sum_{d=1 }^{D}\lambda_{0}\{\tilde{t_d}\} Y_i(\tilde{t_d})$

```{r}
#cumulative function

Funct_lambda2<-function(lambda0,Ts,event){
  #lambda0<- rep(0.1,length(event))
 
  n = length(Ts)
  l =  which(event==0)
  lambda0[l] = 0

  lambda2 = vector()
  for (i in 1:n) {
    vi = observe(Ts[i], Ts)
    lambda2[i] =  sum(lambda0[vi])   
  }
  return(lambda2=lambda2)
}
```

```{r}
lambda0 =  rep(0.1,length(event))
lambda2=  Funct_lambda2(lambda0,Ts,event)
head(lambda2)

```

3) Calculates the conditional probabilities of having $Z=x_j$ ( with the function "Functio_prob")

$\pi_{ij}( \boldsymbol{\theta}^{r})= \frac{   p_{ij} \left[ \lambda_{0}(t_{i}) \exp \left({\boldsymbol{\beta^\top} \boldsymbol{x}_{j}} \right) \right]^{\delta_i} \exp\left(- \Lambda_0(t_{i}) \exp ({\boldsymbol{\beta^\top} \boldsymbol{x}_{j}} )\right)  }{ \sum_{j=1}^{n_{B}}     p_{ij} \left[ \lambda_{0}(t_{i}) \exp \left({\boldsymbol{\beta^\top} \boldsymbol{x}_{j}} \right) \right]^{\delta_i} \exp\left(- \Lambda_0(t_{i}) \exp ({\boldsymbol{\beta^\top} \boldsymbol{x}_{j}} )\right)}$

```{r}
### proba aposteriorie

Functio_prob<-function(beta0,lambda0,lambda2,Ts,event,XB, Q){
  #beta0<- c(0.1,0.1) 
  #lambda0<- rep(0.1,length(event))
  
  X = as.matrix(XB, ncol = p) # covariates

  eXbeta0 = exp(X%*% beta0)
  XeXbeta0 = X*matrix(rep(exp(X%*% beta0),p), ncol = p)

  n = length(Ts)
  m = nrow(XB)
  
#pi(ij)

prob = matrix(0,n,m)
for (i in 1:n) {
  
  numerateur = vector()
  for (j in 1:m) {
    numerateur[j] = Q[i,j] * ( lambda0[i]*eXbeta0[j])^event[i] * exp(-lambda2[i]* eXbeta0[j] )
  }
  
  denominateur = sum(numerateur)
  
  prob[i,] = numerateur/denominateur
}
prob
return( prob = prob)
}
```


```{r}
lambda0 =  rep(0.1,length(event))
beta0 = c(0.1,0.1)
p_f =  Functio_prob(beta0,lambda0,lambda2,Ts,event,XB, Q)
S=apply(p_f,1, sum)
S
prob = p_f  
```


4)Calculation of the baseline risk $(k \geq 1)$ ("Function_lambda0")

   $\lambda_{0}^{(k)}\{\tilde{t_{d}}\} = \frac{1 }{ \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}\pi_{ij}(\boldsymbol{\theta}^{r}) \exp (\boldsymbol{\beta^\top} \boldsymbol{x}_{j}) Y_i(\tilde{t_{d}} )} $


```{r}
################ estimations ###################################

#lambda0
Function_lambda0<-function(prob,beta0,Ts,event,XB){
  
  p = ncol(XB)
  n = length(Ts)
  
  X = as.matrix(XB, ncol = p) # covariates
  eXbeta0 = exp(X%*% beta0)
 
  som1 = prob%*%eXbeta0 # somme sur les j

  lambda0_1 = vector()
  
  for (i in 1:n) {
    
    if(event[i] == 0){lambda0_1[i] = 0
    }else if(event[i] != 0 ){
      
    risq = GetRiskSet (Ts[i], Ts, event)
    nrisk = length(risq)
    
    if(nrisk==1){
      t2R = som1[risq] 
    }else if(nrisk!=1){
      
      t2R = sum( som1[risq] )
    }
    
    lambda0_1[i] = (1/t2R)
  }
  }
  return( lambda0_1 = lambda0_1)
}
```

```{r}
lambda0_1 = Function_lambda0 (prob,beta0,Ts,event,XB)
head(lambda0_1)

```


5) Solving the estimating equation: estimation of beta at iteration $(k \geq 1)$ ("equa_estimate" and "coxph_estimate")


$H_w^{3} (\boldsymbol{\beta}^{(k)})=\sum_{i=1}^{n_A}\delta_{i}\left(\sum_{j=1}^{n_{B}}\pi_{ij}(\boldsymbol{\theta}^{r})\boldsymbol{x}_{j} - \frac{ \sum_{k=1}^{n_{A}} \sum_{j=1}^{n_{B}}   Y_k(\tilde{t_{i}} ) \pi_{kj}(\boldsymbol{\theta}^{r})  \exp\left( \boldsymbol{\beta}^{(r+1)^\top} \boldsymbol{x}_j \right)\boldsymbol{x}_{j} } {  \sum_{k=1}^{n_A}\sum_{j=1}^{n_B}  Y_k(\tilde{t_{i}} )\pi_{kj}(\boldsymbol{\theta}^{r}) \exp(\boldsymbol{\beta}^{(r+1)^\top} \boldsymbol{x}_{j}) }\right)=0$



```{r}
#  estimating equation ########################

equa_estimate <- function(beta,prob,Ts,event,XB) {
  
  p = ncol(XB)
  n = length(Ts)

  X = as.matrix(XB, ncol = p) # covariates
  eXbeta = exp(X%*% beta)
  XeXbeta = X*matrix(rep(exp(X%*% beta),p), ncol = p)
  
  #######les sommes pour Z

  Z = prob%*%X 

  #### les sommes q*exp(beta x) pour tilfe_f
  
  som1 = prob%*%eXbeta

  ###  les sommes pour tilde_g
  
  som2 = prob%*%XeXbeta

  dat1 = cbind(Ts, Z)[which(event==1),]
  
  ## Estimating equation
  s = matrix(0,nrow=nrow(dat1), ncol = p)
  
  for (i in 1:nrow(dat1)) {
    ts = dat1[i, 1]
    Z1R = dat1[i,2:ncol(dat1)]
    
    risk = GetRiskSet(ts, Ts, event)
    nrisk = length(risk)
    
    if(nrisk==1){
      t2R = som1[risk] 
      t3R = matrix(som2, ncol = p)[risk,] 
    }else if(nrisk!=1){
      
      t2R = sum( som1[risk] )
      t3R = colSums(matrix(som2, ncol = p)[risk,] ) }
    
    s[i,] = (Z1R - t3R/t2R)
    
  }
  
  s <- colSums(s)
  
  return(s=s)
  
}
```

```{r}
# ###########solve the equation 

coxph_estimate<- function(beta,prob,Ts,event,XB,beta_ini,maxiter = 50){
  
  f <- function(x){
    equa_estimate  (beta= x,prob,Ts,event,XB)
  }
  fit_manual <- nleqslv( c(0,0),f, method = c("Broyden", "Newton"))
  beta0 <- fit_manual$x
  iterations <- fit_manual$iter
  converge <- as.numeric(( iterations< maxiter)& (fit_manual$scalex==1) )
  
  #xstart <- matrix(rnorm(20,0,1), ncol = 2)
  #Zero <-  searchZeros(xstart,f)
  #fit_manual <- multiroot(f,start = rep(0,p), maxiter = maxiter)
 # beta0 <- fit_manual$root
  #iterations <- fit_manual$iter
#converge <- as.numeric((fit_manual$iter < maxiter)& !is.nan(fit_manual$estim.precis) & !is.na(fit_manual$estim.precis) & (fit_manual$estim.precis<1e-6))
  
  return(list( beta0 = beta0, converge = converge, iterations = iterations) )
}
```


```{r}
beta_ini=c(0,0)
H_f = coxph_estimate (beta,prob,Ts,event,XB,beta_ini, maxiter = 50)
H_f$beta0
H_f$converge
H_f$iterations

```
Conditions:( relative errors) ("Func_itteration")


 $\mid \mid \theta_0^{(k+1)} - \theta_0^{(k)} \mid\mid< 10^{-6}$ 
    

```{r}
### iterrations ###########################################

#valeurs initials

Func_itteration<-function(beta,beta0,lambda0,Ts,event,XB, Q,tol= 1e-6, maxits = 500){
  
  p = ncol(XB)
  n = length(Ts)
  
  it = 0
  converge = FALSE
  
while ((!converge) & (it < maxits)){ 

  lambda0.old = lambda0
  beta0.old = beta0
  beta_ini = beta0.old
  
  #expectation  
  
  lambda2 = Funct_lambda2(lambda0.old,Ts,event)

  prob = Functio_prob(beta0.old,lambda0.old,lambda2,Ts,event,XB, Q)

  
  #maximization
  lambda0 = Function_lambda0 (prob,beta0, Ts,event,XB)
  estime = coxph_estimate (beta,prob,Ts,event,XB,beta_ini, maxiter = 50)
  beta0 = estime$beta0

  converge = (sum ((beta0.old -beta0))^2 + sum ( (lambda0.old - lambda0)^2 ) ) < tol  
           

  if (it == maxits) {
    cat("WARNING! NOT CONVERGENT!", "\n")
    converge = FALSE
  }else if(is.na(beta0[1]) & is.na(beta0[2])){
    cat("WARNING! beta0 NOT AVAILABLE!", "\n")
    converge = FALSE
  }
  
   it = it + 1

}
  
return(list(beta0=beta0, lambda0=lambda0, prob=prob, converge= converge))
}

```


```{r}
#############################################################
#method 4
p = 2
m    = 500
n    = round ( (80*m)/100)
beta = c(0.5,-0.5)
C = c(0.8,0.1,0.1)
beta_ini=c(0,0)
      surv_data = Generate_data(m,n,beta)
      mf = Matrix_function(n,m,C,beta,surv_data)
      data_true = mf$data_true
      data_naive = mf$data_naive

      Ts = mf$data_true$Time
      event = mf$data_true$delta
      Q = mf$Q
      XB = mf$XB
      Z = as.matrix(data_naive[,3:(p+2)]) 
      
      lambda0 =  rep(0.1,length(event))
      beta0 = c(0.1,0.1)

  fit_estimate = Func_itteration(beta,beta0,lambda0,Ts,event,XB, Q,tol= 1e-6, maxits = 500)
  coef_estimate = fit_estimate$beta0
  lambda0_estimate = fit_estimate$lambda0
  converge_estimate = fit_estimate$converge

  # Theoretical estimating equation for true and naive data
  fit_true = coxph(Surv(Time,delta)~.,data = data_true)
  coef_true = as.vector(fit_true$coefficients)
  var_true = diag(fit_true$var)

  coef_true
  coef_estimate
```

# equation naive
```{r}
equa_naive <- function(beta,Ts,event, Z) {
  
  p=ncol(Z)
  ezbeta = exp(Z%*% beta)
  zezbeta = Z*matrix(rep(exp(Z%*% beta),p), ncol = p) 
  
  n = length(Ts)
  
  s = 0
  for(i in 1:n){ 
    
#Yi<- as.numeric( Ts >= Ts[i])# at risk
    Yi = as.numeric( (Ts == Ts[i] ) | (Ts > Ts[i]))# at risk
    risq = which(Yi==1)
    
    if(length(risq)==1){
      
      num_naive = zezbeta[risq,]
      denum_naive = ezbeta[risq] 
      
    }else if (length(risq)!=1){
      
      num_naive = colSums( zezbeta[risq,])
      denum_naive = sum(ezbeta[risq])  }  
    
    s = s+ event[i]* ( Z[i,]- num_naive/denum_naive)
  }
  s
  return(H_naive = s) ##naive estimator 
}
```


```{r}
# solve the naive equation 

coxph_equa_naive <- function(Ts,event, Z, maxiter = 20){
  p=ncol(Z)
  
  f <- function(x){
    equa_naive (beta=x,Ts,event, Z)
  }
  
  fit_manual = multiroot(f,start = rep(0,p), maxiter = maxiter)
  beta = fit_manual$root
  iterations = fit_manual$iter
  converge = as.numeric((fit_manual$iter < maxiter)& !is.nan(fit_manual$estim.precis) & !is.na(fit_manual$estim.precis) & (fit_manual$estim.precis<1e-6))
  
  return(list(coef = beta, converge = converge, iterations = iterations))
}

```

```{r}

```

```{r}
  
  fit_naive2 = coxph(Surv(Time,delta)~.,data = data_naive)
  coef_naive2 = as.vector(fit_naive2$coefficients)
  var_naive2 = diag(fit_naive2$var)

  fit_naive = coxph_equa_naive(Ts,event, Z, maxiter = 20)
  coef_naive = fit_naive$coef
  converge_naive = fit_naive$converge
  
  coef_true
  coef_naive 
  coef_naive2
 
  
```

Method 2 presentation: Weighted-sum-method 1

- Assuming that the matching probabilities matrix $Q$ is known, 

- First, let us consider the surogate variables $\tilde{Z_i}$  as being the weighted average of all the covariables $X_j$ such as,

$\tilde{Z_i}=\sum_{j=1}^{n_B} p_{ij}X_j$, $ \forall j \in XB$,

- we estimate the beta parameter by solving the first estimating equation "equa_W_sum1" using Nweton's numerical method in the function "coxph_equa_W_sum1".


```{r}
#################

equa_W_sum1 <- function(beta,Ts,event,XB, Q) {
  
  p = ncol(XB)
  X = as.matrix(XB, ncol = p) # covariates
   
  eXbeta = exp(X%*% beta)
  XeXbeta = X*matrix(rep(exp(X%*% beta),p), ncol = p)
  
  ##les sommes pour Z
  Z = Q%*%X
  
  ## les sommes q*exp(beta x) pour tilfe_f
  som1 =  Q%*%eXbeta
  
  ##  les sommes pour tilde_g
  som2 = Q%*%XeXbeta
  
  dat1 = cbind(Ts,Z)[which(event==1),] #new  data
  
  ## Estimating equation
  s = matrix(0,nrow=nrow(dat1), ncol = p)
  
  for (i in 1:nrow(dat1)) {
    ts = dat1[i, 1]
    Z1R = dat1[i,2:ncol(dat1)]
    
    risk = GetRiskSet(ts, Ts, event)
    nrisk = length(risk)
    if(nrisk==1){
      t2R = som1[risk] 
      t3R = matrix(som2, ncol = p)[risk,] 
      
    }else if(nrisk!=1){
      
      t2R = sum(som1[risk] )
      t3R = colSums(matrix(som2, ncol = p)[risk,] ) }
    
    s[i,] = (Z1R - (t3R/t2R))
    
  }
  
  s = colSums(s)
  
  return(s=s)
  
}
```

```{r}
# ###########solve the equations 
coxph_w_sum1 <- function(Ts,event,XB, Q, maxiter = 20){
  
  f <- function(x){
    equa_W_sum1  (beta=x,Ts,event,XB, Q)
  }
  
  fit_manual = multiroot(f,start = rep(0,p), maxiter = maxiter)
  beta = fit_manual$root
  iterations = fit_manual$iter
  converge = as.numeric((fit_manual$iter < maxiter)& !is.nan(fit_manual$estim.precis) & !is.na(fit_manual$estim.precis) & (fit_manual$estim.precis<1e-6))
  
  return(list(coef = beta, converge = converge,iterations = iterations))
}
```



```{r}
#############################

  #second method
  fit_w_sum1 = coxph_w_sum1(Ts,event,XB, Q,maxiter = 20)
  coef_w_sum1 = fit_w_sum1$coef
  converge_w_sum1 = fit_w_sum1$converge
  
  coef_true
  coef_w_sum1
```

Method 3 presentation: Weighted-sum-method 2

- Consider the same $Q$ matrix

- In order to reduce the bias, we consider the case where the surogate variable $Z_i$ is a weighted average of  only the two covariates with the highest matching probabilities $ p_{ij_1}$ and $ p_{ij_2}$ such as,
 
 $Z_i= p_{ij_1}Z_{j_1} + p_{ij_2}Z_{j_2}$,  with $j_1 , j_2 \in XB$.
 
 we estimate the beta parameter by solving the estimating equation "equa_W_sum2" using Nweton's numerical method in "coxph_equa_W_sum2".


```{r}
#################

equa_W_sum2 <- function(beta,Ts,event,XB, Q) {
  
  p = ncol(XB)
  n = length(Ts)
  X = as.matrix(XB, ncol = p) #Only covariates
  
  eXbeta = exp(X%*% beta)
  XeXbeta = X*matrix(rep(exp(X%*% beta),p), ncol = p)
  
  
  
  # somme des deux valeurs max
  Z = matrix(0, nrow = n, ncol = p)
  som1 = vector()                  # denominator
  som2 = matrix(0, nrow = n, ncol = p) #numerator
  
   for (i in 1:n) {
    qi = Q[i,]
    i1 = which(qi == max(qi))
    l1 = max(qi)
    qi[i1] = 0
    i2 = which(qi == max(qi))
    if(length(i2)!= 1){i2 = i2[1]}
    qi[i1] = l1
    
    qix = (qi*X)[c(i1,i2),]
    Z[i,] = colSums(qix)
    
  #### les sommes q*exp(beta x) pour tilfe_f   
    qixbeta = (qi*eXbeta)[c(i1,i2),]
    som1[i] = sum( qixbeta)
    
  ###  les sommes pour tilde_g   
    qiXeXbeta = (qi*XeXbeta)[c(i1,i2),]
    som2[i,] = colSums(qiXeXbeta)
  }  
 
  
  dat1 = cbind(Ts,Z)[which(event==1),]
  ## Estimating equation
  s = matrix(0,nrow=nrow(dat1), ncol = p)
  for (i in 1:nrow(dat1)) {
    ts = dat1[i, 1]
    Z1R = dat1[i,2:ncol(dat1)]
    
    risk = GetRiskSet(ts, Ts, event)
    nrisk = length(risk)
    if(nrisk==1){
      t2R = som1[risk] 
      t3R = matrix(som2, ncol = p)[risk,] 
      
    }else if(nrisk!=1){
      
      t2R = sum(som1[risk] )
      t3R = colSums(matrix(som2, ncol = p)[risk,] ) }
    
    s[i,] = (Z1R - (t3R/t2R))
    
  }
  
  s = colSums(s)
  return(H_w_sum2 = s)  
}
```

```{r}
# ###########solve the equations 


coxph_w_sum2 <- function(Ts,event,XB, Q, maxiter = 20){
  p= ncol(XB)
  
  f <- function(x){
    equa_W_sum2  (beta=x,Ts,event,XB, Q)
  }
  
  fit_manual <- multiroot(f,start = rep(0,p), maxiter = maxiter)
  beta <- fit_manual$root
  iterations <- fit_manual$iter
  converge <- as.numeric((fit_manual$iter < maxiter)& !is.nan(fit_manual$estim.precis) & !is.na(fit_manual$estim.precis) & (fit_manual$estim.precis<1e-6))
  
  return(list(coef = beta, converge = converge, iterations=iterations))
}
```


```{r}
#############################
  ### method 3
  fit_w_sum2 = coxph_w_sum2(Ts,event,XB, Q, maxiter = 20)
  coef_w_sum2 = fit_w_sum2$coef
  converge_w_sum2 = fit_w_sum2$converge
 
  coef_true
  coef_w_sum2
```




```{r}
################### monte carlos

estimates_survival<- function(nsim,n,m,C,p,beta,Q,lambda0,beta0){
  
  
  coef_true_s = matrix(0,nrow = nsim, ncol = p)
    
  coef_naive_s = matrix(0,nrow = nsim, ncol = p)
  converge_naive = vector()

  coef_w_sum1_s = matrix(0,nrow = nsim, ncol = p)
  converge_w_sum1 = vector()

  coef_w_sum2_s = matrix(0,nrow = nsim, ncol = p)
  converge_w_sum2 = vector()

  coef_estimate_s = matrix(0,nrow = nsim, ncol = p)
  converge_estimate = vector()
  
 
  for (i in 1:nsim){
    
      
      surv_data = Generate_data(m,n,beta)
      mf = Matrix_function(n,m,C,beta,surv_data)
      data_true = mf$data_true
      data_naive = mf$data_naive

      Ts = mf$data_true$Time
      event = mf$data_true$delta
      Q = mf$Q
      XB = mf$XB
      Z = as.matrix(data_naive[,3:(p+2)]) 
      
      lambda0 =  rep(0.1,length(event))
      beta0 = c(0.1,0.1)

     # prob =  Functio_prob(beta0,lambda0,Ts,event,XB, Q)
      
      # Theoretical estimating equation for true and naive data
      fit_true = coxph(Surv(Time,delta)~.,data = data_true)
      coef_true_s[i,] = as.vector(fit_true$coefficients)
      
      fit_naive = coxph_equa_naive(Ts,event, Z, maxiter = 20)
      coef_naive_s[i,] = fit_naive$coef
      converge_naive[i] = fit_naive$converge
      
      fit_w_sum1 = coxph_w_sum1(Ts,event,XB, Q,maxiter = 20)
      coef_w_sum1_s[i,] = fit_w_sum1$coef
      converge_w_sum1[i] = fit_w_sum1$converge
      
      fit_w_sum2 = coxph_w_sum2(Ts,event,XB, Q, maxiter = 20)
      coef_w_sum2_s[i,] = fit_w_sum2$coef
      converge_w_sum2[i] = fit_w_sum2$converge
 
      fit_estimate = Func_itteration(beta,beta0,lambda0,Ts,event,XB, Q,tol= 1e-6,maxits = 500)
      coef_estimate_s[i,] = fit_estimate$beta0
      converge_estimate[i] = as.numeric(fit_estimate$converge)
      

    }
  
  return(list( coef_true_s1 = coef_true_s[,1], coef_true_s2 = coef_true_s[,2],
   coef_naive_s1 = coef_naive_s[,1], coef_naive_s2 = coef_naive_s[,2], coef_w_sum1_s1 = coef_w_sum1_s[,1], 
   coef_w_sum1_s2 = coef_w_sum1_s[,2], coef_w_sum2_s1 = coef_w_sum2_s[,1], coef_w_sum2_s2 = coef_w_sum2_s[,2],  
coef_estimate_s1 = coef_estimate_s[,1], coef_estimate_s2 = coef_estimate_s[,2], converge_naive = converge_naive, 
converge_w_sum1= converge_w_sum1,converge_w_sum2 = converge_w_sum2, converge_estimate = converge_estimate))  } 
######################################
```

```{r}
p = 2
m    = 500
n    = round((80*m)/100)
beta = c(0.5,-0.5)
C = c(0.8,0.1,0.1)
beta_ini=c(0,0)
nsim=3

r_m = estimates_survival (nsim,n,m,C,p,beta,Q,lambda0,beta0)
R_T =cbind( mean ( r_m$coef_true_s1), mean(r_m$coef_true_s2))
R_N = cbind( mean ( r_m$coef_naive_s1), mean(r_m$coef_naive_s2 ))
R_2 = cbind( mean ( r_m$coef_w_sum1_s1 ), mean(r_m$coef_w_sum1_s2 ))
R_3 = cbind( mean ( r_m$coef_w_sum2_s1 ), mean(r_m$coef_w_sum2_s2 ))
R_4 = cbind( mean ( r_m$coef_estimate_s1 ), mean(r_m$coef_estimate_s2 ))
```

```{r}
rbind(R_T, R_N ,R_2, R_3, R_4)

```


Simulation with some scenarios.
```{r}

C_sample = cbind(c(0.6,0.2,0.2),c(0.7,0.15,0.15),c(0.8,0.1,0.1))
m = c(200,500,800)
n = round((80*m)/100)
p = 2
nsim = 500
beta = c(0.5,-0.5)
beta_ini=c(0,0)
#lambda0 = rep(0.5,n)
#beta0 = c(0.1,0.1) 

##############

scenarios = NULL
for (i in 1:ncol(C_sample)){
  scenarios = rbind(scenarios,c(nsim,n[2],m[2],C_sample[,i]))
}
for (i in 4:6){
  scenarios = rbind(scenarios,c(nsim,n[i-3],m[i-3],C_sample[,3]))
}

colnames(scenarios) = c("nsim","n","m","Prob_1","Prob_2","Prob_3")

scenarios=data.frame(scenarios)
```

```{r}
scenarios
```

Results

```{r}

setwd("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu")

for (i in (1:nrow(scenarios))){
  
  nsim=scenarios[i,1]
  n=scenarios[i,2]
  m=scenarios[i,3]
  C=vector()
  C[1]=scenarios[i,4]
  C[2]=scenarios[i,5]
  C[3]=scenarios[i,6]
  
  lambda0 = rep(0.1,n)
  beta0 = c(0.1,0.1) 
  
  results_sample = estimates_survival(nsim,n,m,C,p,beta,Q,lambda0,beta0) #monte carlos

 #filename = paste0("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu/Results/","nsim=",nsim," n = ",n," m = #",m,"  Prob_1 = ",C[1]," Prob_2 = ",C[2]," Prob_3 = ", C[3],".Rdata")
  filename = paste0("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu/Results/","scenario=",i,
                    "_m=",m, "_prob1=", C[1], ".Rdata")

  save(results_sample,file = filename)
  
}

```

```{r}
# one file
setwd("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu/Results/")
load("scenario=1_m=1000_prob1=0.6.Rdata")
results_sample = results_sample
```


```{r}

# list of files to load

############results


setwd("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu")
#source("8_scenarios.R")

# list of files to load
#library(parallel) 

file_list <- c("scenario=1_m=500_prob1=0.6.Rdata", 
               "scenario=2_m=500_prob1=0.7.Rdata",  
               "scenario=3_m=500_prob1=0.8.Rdata",
               "scenario=4_m=200_prob1=0.8.Rdata",
               "scenario=5_m=500_prob1=0.8.Rdata",
               "scenario=6_m=800_prob1=0.8.Rdata") 


r = nrow(scenarios) 

funct_tableau <- function(file_list,r,p){
  
  beta_true = matrix(0,r,p )
  beta_naive = matrix(0,r,p )
  beta_w_sum1 = matrix(0,r,p )
  beta_w_sum2 = matrix(0,r,p )
  beta_estimate = matrix(0,r,p )
  
  L_naive = vector()  # number of convergences
  L_w_sum1 = vector()
  L_w_sum2 = vector()
  L_estimate = vector()
  
  Sd_true = matrix(0,r,p ) # standard deviation
  Sd_naive = matrix(0,r,p )
  Sd_sum1 = matrix(0,r,p )
  Sd_sum2 = matrix(0,r,p )
  Sd_estimate = matrix(0,r,p )
  
  rmse_true = matrix(0,r,p ) # mean squart error
  rmse_naive = matrix(0,r,p )
  rmse_sum1 = matrix(0,r,p )
  rmse_sum2 = matrix(0,r,p )
  rmse_estimate = matrix(0,r,p )
  
  for (i in 1:length(file_list)) {
    
    setwd("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu/Results")
    load(file_list[i])
    results_sample = results_sample
    
    beta_true[i,] = c (mean(results_sample$coef_true_s1), mean(results_sample$coef_true_s2))
    beta_naive[i,] = rbind(  mean(results_sample$coef_naive_s1), mean(results_sample$coef_naive_s2))
    beta_w_sum1[i,] = rbind( mean(results_sample$coef_w_sum1_s1), mean(results_sample$coef_w_sum1_s2))
    beta_w_sum2[i,] = rbind( mean(results_sample$coef_w_sum2_s1), mean(results_sample$coef_w_sum2_s2))
    beta_estimate[i,] = rbind( mean(results_sample$coef_estimate_s1 ), mean(results_sample$coef_estimate_s2 ))
    
    
    L_naive[i] = length(which(results_sample$converge_naive==0))
    L_w_sum1[i] = length(which(results_sample$converge_w_sum1==0))
    L_w_sum2[i] = length( which(results_sample$converge_w_sum2==0))
    L_estimate[i] = length(which(results_sample$converge_estimate==0))
    
    
    Sd_true[i,] = rbind( sd(results_sample$coef_true_s1), sd(results_sample$coef_true_s2))
    Sd_naive[i,] = rbind( sd( results_sample$coef_naive_s1) , sd( results_sample$coef_naive_s2) )
    Sd_sum1[i,] = rbind( sd(results_sample$coef_w_sum1_s1), sd(results_sample$coef_w_sum1_s2))
    Sd_sum2 [i,] = rbind( sd(results_sample$coef_w_sum2_s1),sd(results_sample$coef_w_sum2_s2))
    Sd_estimate[i,] = rbind( sd(results_sample$coef_estimate_s1),sd(results_sample$coef_estimate_s2))
    
    
    rmse_true[i,] = rbind (sqrt( mean((results_sample$coef_true_s1- beta[1])^2)),
                           sqrt( mean((results_sample$coef_true_s2+  beta[2])^2)))
    
    rmse_naive[i,] = rbind( sqrt( mean((results_sample$coef_naive_s1- beta[1])^2)),
                            sqrt( mean((results_sample$coef_naive_s2+  beta[2])^2)))
    
    rmse_sum1[i,] = rbind( sqrt( mean((results_sample$coef_w_sum1_s1- beta[1])^2)),
                           sqrt( mean((results_sample$coef_w_sum1_s2- beta[2])^2)))
    
    rmse_sum2[i,] = rbind( sqrt( mean((results_sample$coef_w_sum2_s1- beta[1])^2)),
                           sqrt( mean((results_sample$coef_w_sum2_s2- beta[2])^2)))
    
    rmse_estimate[i,] = rbind( sqrt( mean((results_sample$coef_estimate_s1 -  beta[1])^2)),
                               sqrt( mean((results_sample$coef_estimate_s2 -  beta[2])^2)))
    
    
  }
  
  return( list( beta_true = beta_true, beta_naive=beta_naive,beta_w_sum1=beta_w_sum1,
   beta_w_sum2=beta_w_sum2, beta_estimate=beta_estimate,L_naive =L_naive,L_w_sum1=L_w_sum1,L_w_sum2=L_w_sum2,L_estimate=L_estimate, Sd_true= Sd_true,Sd_naive=Sd_naive ,Sd_sum1=Sd_sum1,  Sd_sum2=Sd_sum2, Sd_estimate =Sd_estimate, rmse_true = rmse_true, rmse_naive = rmse_naive,rmse_sum1 = rmse_sum1, rmse_sum2 = rmse_sum2, rmse_estimate = rmse_estimate))
  
}
#####################

Results_scenario = as.data.frame(funct_tableau(file_list,r,p))
rownames(Results_scenario )= paste0( "Scenario", 1:nrow(Results_scenario))


```


```{r}
r = nrow(scenarios) 
p = 2
 Results_scenario = as.data.frame(funct_tableau(file_list,r,p))
```

```{r}


### boxplots


scenarios_boxplot=scenarios # the last scenario 6

library(ggplot2)
library(gridExtra)

levelsvec = numeric(0)


file_list <- c("scenario=1_m=500_prob1=0.6.Rdata", 
               "scenario=2_m=500_prob1=0.7.Rdata",  
               "scenario=3_m=500_prob1=0.8.Rdata",
               "scenario=4_m=200_prob1=0.8.Rdata",
               "scenario=5_m=500_prob1=0.8.Rdata",
               "scenario=6_m=800_prob1=0.8.Rdata") 

for (i in (1:nrow(scenarios_boxplot))){
  nsim = scenarios_boxplot[i,1]
  n = scenarios_boxplot[i,2]
  m = scenarios_boxplot[i,3]
  Prob1 = scenarios_boxplot[i,4]
  Prob2 = scenarios_boxplot[i,5]
  Prob3 = scenarios_boxplot[i,6]

  Prob_sample = paste0("(",Prob1,",",Prob2,",",Prob3,")") 
  
  setwd("C:/Users/fchezeut/Documents/GitHub/Cox_matched_data/code_Q_connu/Results")
  load(file_list[i])
  results_sample = results_sample
  
  levelsvec = c(levelsvec,Prob_sample)

  boxplot_beta0_true = cbind(rep(1,times=nsim),
                               rep(Prob_sample,times=nsim),
                               results_sample$coef_true_s1,
                               results_sample$coef_true_s2)
  
  boxplot_beta0_naive = cbind(rep(2,times=nsim),
                      rep(Prob_sample,times=nsim),
                      results_sample$coef_naive_s1,
                      results_sample$coef_naive_s2)
  
  boxplot_beta0_w_sum1 = cbind(rep(3,times=nsim),
                      rep(Prob_sample,times=nsim),
                      results_sample$coef_w_sum1_s1,
                      results_sample$coef_w_sum1_s2)
  
  boxplot_beta0_w_sum2 = cbind(rep(4,times=nsim),
                                rep(Prob_sample,times=nsim),
                                results_sample$coef_w_sum2_s1,
                                results_sample$coef_w_sum2_s2)
  
  boxplot_beta0_estimate = cbind(rep(5,times=nsim),
                                rep(Prob_sample,times=nsim),
                                results_sample$coef_estimate_s1,
                                results_sample$coef_estimate_s2)
  
  boxplot_beta0 = rbind(boxplot_beta0_true,boxplot_beta0_naive,boxplot_beta0_w_sum1,
                        boxplot_beta0_w_sum2,boxplot_beta0_estimate )
  
  
  titlename = paste0("n=",unique(scenarios_boxplot[i,2]),
                   ", ","m=",unique(scenarios_boxplot[i,3]))

colnames(boxplot_beta0) = c("Estimator", "Prob","beta0[1]","beta0[2]")

boxplot_beta0 = data.frame(boxplot_beta0)

boxplot_beta0[,1] = factor(boxplot_beta0[,1],
      levels = 1:5, labels = c("True"," Naive", "w_sum1", "w_sum2", "EM"))

boxplot_beta0[,2] = factor(boxplot_beta0[,2],
                         levels = levelsvec,labels = levelsvec)

boxplot_beta0[,3] = as.numeric(boxplot_beta0[,3] )
boxplot_beta0[,4] = as.numeric(boxplot_beta0[,4] )
# Boxplots for beta0

beta0_1 = ggplot(boxplot_beta0,aes(x=Prob,y= beta0.1.,fill=Estimator))+
  geom_boxplot()+
  xlab("Q_matrix")+
  ylab(quote(hat(beta)[0]))+
  geom_hline(yintercept=0.5,lty=1,col="orange")+
  ggtitle(titlename)


beta0_2 = ggplot(boxplot_beta0,aes(x=Prob,y= beta0.2.,fill=Estimator))+
  geom_boxplot()+
  xlab("Q_matrix")+
  ylab(quote(hat(beta)[0]))+
  geom_hline(yintercept=-0.5,lty=1,col="orange")+
  ggtitle(titlename)

print(beta0_1)
print(beta0_2)

  
}

```

```{r}

titlename = paste0("n=",unique(scenarios_boxplot[,2]),
                   ", ","m=",unique(scenarios_boxplot[,3]))

colnames(boxplot_beta0) = c("Estimator", "Prob","beta0[1]","beta0[2]")

boxplot_beta0 = data.frame(boxplot_beta0)

boxplot_beta0[,1] = factor(boxplot_beta0[,1],
      levels = 1:5, labels = c("True"," Naive", "w_sum1", "w_sum2", "EM"))

boxplot_beta0[,2] = factor(boxplot_beta0[,2],
                         levels = levelsvec,labels = levelsvec)

boxplot_beta0[,3] = as.numeric(boxplot_beta0[,3] )
boxplot_beta0[,4] = as.numeric(boxplot_beta0[,4] )
# Boxplots for beta0

beta0_1 = ggplot(boxplot_beta0,aes(x=Prob,y= beta0.1.,fill=Estimator))+
  geom_boxplot()+
  xlab("Q_matrix")+
  ylab(quote(hat(beta)[0]))+
  geom_hline(yintercept=0.5,lty=1,col="orange")+
  ggtitle(titlename)
beta0_1

beta0_2 = ggplot(boxplot_beta0,aes(x=Prob,y= beta0.2.,fill=Estimator))+
  geom_boxplot()+
  xlab("Q_matrix")+
  ylab(quote(hat(beta)[0]))+
  geom_hline(yintercept=-0.5,lty=1,col="orange")+
  ggtitle(titlename)
beta0_2

```



boxplot
```{r}


library(xtable)
s = 5*r
tb = matrix(0, nrow = s, ncol = 11) # table of summary 
for (i in s) {
  tb[,i] = c(   )
}

tb <- xtable(Base_plot,digits=3,caption="Résultats sur les profils-lignes pour les 3 premiers axes factoriels")


Base_plot<- as.data.frame(results_sample[1:10] )
colnames(Base_plot)=c( "beta_true_1", "beta_true_2","beta_naive_1","beta_naive_2", "beta_w_sum1_1","beta_w_sum1_2","beta_w_sum2_1", "beta_w_sum2_2", "beta_estimate_1", "beta_estimate_2")


print(tb,include.rownames=TRUE,  sanitize.text.function = function(x){x},tabular.environment='longtable', hline.after = c(-1,0),floating=FALSE,size="\\scriptsize")

results_sample_plot <- boxplot(Base_plot, col=c("gray","gray","red","red", "green","green","blue","blue","orange", "orange" ), boxwex=2)



####################
```



## 



```{r}

```

## 

```{r}


```
